---
title: "Association Rules"
author: "Che Cobb, Andy Heroy, Carson Drake, David Josephs"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: TRUE
    toc_float:
      collapsed: true
      smooth_scroll: false
    theme: paper
    df_print: paged
    keep_md: TRUE

---

```{r setup, echo = F}
if (knitr::is_latex_output()) {
	knitr::opts_chunk$set(dev = "tikz")
}
knitr::opts_knit$set(cache =T, autodep = T)
knitr::opts_chunk$set(warning = F)
knitr::opts_chunk$set(message = F)
knitr::opts_chunk$set(tidy = T, tidy.opts = list(comment = F))
knitr::opts_chunk$set(fig.align = "center")
knitr::opts_chunk$set(comment = '#>')
knitr::opts_chunk$set(fig.path = 'fig/')
library(knitr)
```

```{r, echo = F}
read_chunk('../analysis/Lab3Start.R')
```

# Setup

First, we need to load in the necessary libraries:

```{r libs}
```

Then we need to load our dataset, as well as examine it a bit

```{r datadef}
```

Next lets get rid of some columns we are not going to use/that has some issues, and then lets break down our numeric variables into factors using `cut`

```{r cleanup}
```

Next, we convert the data to an object with class `transactions`, after viewing it again

```{r transact}
```

# Exploration

Before we begin with our analysis, lets check out the rule frequencies within the dataset. We are looking for rules with `support >= .2`

```{r rulefreq}
```

# Rule mining

Next, lets mine some rules with the apriori algorithm, and then clean up redundant rules. We are still sorting out what to set the minsupp and minconf to.

```{r rulemine}
```

## Rule quality and inspection

Next, let us inspect the rules, and examine their quality

```{r quality}
```

### Plots

First lets view a scatterplot of our rules

```{r scatterplot}
```

Next lets look at  a balloon plot

```{r baloonplot}
```

Parallel plot

```{r plplot}
```

Two key plot

```{r kplot}
```

grouped plot

```{r gplot}
```

